The \dealii Library, Version 9.5

### 2023-03-19 9.5/paper.tex
Copy files -> 9.5
Add sections

\hypersetup{
  pdfauthor={
    Daniel Arndt,
    Wolfgang Bangerth,
    Marco Feder,
    Marc Fehling,
    Rene Gassm\"{o}ller,
    Timo Heister,
    Luca Heltai,
    Martin Kronbichler,
    Matthias Maier,
    Peter Munch,
    Jean-Paul Pelteret,
    Simon Sticko,
    Bruno Turcksin,
    David Wells
  },
  pdftitle={The deal.II Library, Version 9.4, 2022},
}

\title{The \dealii Library, Version 9.4}

@@

   \author[4]{Marco~Feder}
\affil[4]{SISSA,
   International School for Advanced Studies,
   Via Bonomea 265,
   34136, Trieste, Italy.
   {\texttt{marco.feder/luca.heltai@sissa.it}}}

@@

\author[5]{Rene~Gassm{\"o}ller}
\affil[5]{Department of Geological Sciences,
   University of Florida,
   1843 Stadium Road,
   Gainesville, FL, 32611, USA.
  {\texttt{rene.gassmoeller@ufl.edu}}}

\author[6]{Timo~Heister}
 \affil[6]{School of Mathematical and Statistical Sciences,
   Clemson University,
   Clemson, SC, 29634, USA
   {\texttt{heister@clemson.edu}}}

\author[4]{Luca~Heltai}

 \author[7,8]{Martin~Kronbichler}
 \affil[7]{Department of Information Technology,
   Uppsala University,
   Box 337, 751\,05 Uppsala, Sweden.
   {\texttt{martin.kronbichler/simon.sticko@it.uu.se}}}
 \affil[8]{Institute of Mathematics,
   University of Augsburg,
   Universit\"atsstr.~12a, 86159 Augsburg, Germany.
   {\texttt{martin.kronbichler@uni-a.de}}}

\author[9]{Matthias~Maier}
\affil[9]{Department of Mathematics,
  Texas A\&M University,
  3368 TAMU,
  College Station, TX 77845, USA.
  {\texttt{maier@math.tamu.edu}}}

\author[8,10]{Peter Munch}
 \affil[10]{Institute of Material Systems Modeling,
 Helmholtz-Zentrum Hereon,
 Max-Planck-Str. 1, 21502 Geesthacht, Germany.
   {\texttt{peter.muench@hereon.de}}}

\author[11]{Jean-Paul~Pelteret}
\affil[11]{Independent researcher.
{\texttt{jppelteret@gmail.com}}}

\author[7,12]{Simon~Sticko}
\affil[12]{Department of Mathematics and Mathematical Statistics,
   Umeå University,
   SE-90187 Umeå, Sweden}

@@

\author[13]{David Wells}
\affil[13]{Department of Mathematics, University of North Carolina,
  Chapel Hill, NC 27516, USA.
  {\texttt{drwells@email.unc.edu}}}

@@

\begin{abstract}
  This paper provides an overview of the new features of the finite element
  library \dealii, version 9.4.
\end{abstract}

@@

\dealii version 9.4.0 was released June 24, 2022.
This paper provides an
overview of the new features of this release and serves as a citable
reference for the \dealii software library version 9.4. \dealii is an
object-oriented finite element library used around the world in the
development of finite element solvers. It is available for free under the
GNU Lesser General Public License (LGPL). Downloads are available at
\url{https://www.dealii.org/} and \url{https://github.com/dealii/dealii}.

The major changes of this release are:
%
\begin{itemize}
  \item Advances in simplex- and mixed-mesh support (see Section~\ref{sec:simplex});
  \item Repartitioning of distributed meshes (see Section~\ref{sec:repartitioning});
  \item Advances in matrix-free infrastructure (see Section~\ref{sec:mf});
  \item Advances in multigrid infrastructure (see Section~\ref{sec:multigrid});
  \item CutFEM support (see Section~\ref{sec:cut});
  \item Experimental integration of the Computational Geometry Algorithms Library (CGAL) (see Section~\ref{sec:cgalwrappers});
  \item Performance improvements in the particle infrastructure (see Section~\ref{sec:particles});
  \item Support for large MPI buffers and parallel I/O (see Section~\ref{sec:large-mpi});
  \item Improvements to unstructured communication (see Section~\ref{sec:CA});
  \item Three new tutorial programs and one new code gallery program (see Section~\ref{subsec:steps}).
\end{itemize}
%

While all of these major changes are discussed in detail in
Section~\ref{sec:major}, there
are a number of other noteworthy changes in the current \dealii release,
which we briefly outline in the remainder of this section:
%
\begin{itemize}
%  \item \texttt{AffineConstraints::make\_consistent\_in\_parallel()} allows
%  to make constraints consistent in parallel.
  \item The \texttt{DataOutResample} class interpolates values defined on one
  triangulation onto a second, potentially unrelated triangulation.
  By using this class, one can output the result obtained on an
  unstructured mesh on a structured one (which might facilitate a more
  memory-efficient storage format, for example, if this second
  triangulation is a uniformly refined rectangle or box), or one can create a slice in 3D.

\item The new member function \texttt{find\_point\_owner\_rank()} of
        \texttt{parallel\allowbreak ::distributed::\allowbreak Tri-\allowbreak angulation} allows one to find the MPI
  ranks of cells containing specified points.
  It is communication-free and leverages the functionality of \pfrst (>v.2.2).
  Its algorithm is described in \cite{burstedde2020parallel}. This information will
  enable efficient construction of the
  communication pattern used in the class \texttt{Utilities::\allowbreak MPI::\allowbreak RemotePointEvaluation}. Furthermore, this function could be used in the future to allow
  particle simulations in which particle movement is not
  limited by CFL conditions, as done in \cite{mirzadeh2016parallel}.

\item The new function
  \texttt{GridGenerator::pipe\_junction()}
  generates a triangulation of three cone-shaped pipes that cross at a bifurcation point in any possible configuration.
  A manifold description is applied to the boundary, which can be extended into the volume via transfinite interpolation \cite{Gordon82} using the \texttt{TransfiniteInterpolationManifold} class \cite{dealII90}.
  \item A new DoF renumbering function \texttt{DoFRenumbering::support\_point\_wise()} which groups together
  shape functions by their support point. This functionality is useful in developing nodal schemes since, e.g., the $x$, $y$, and $z$
  components at a point will be consecutive in the solution vector. It also improves interoperability with external libraries which expect
  data in this format.
  \item The \texttt{FEInterfaceValues} class, which computes common quantities at the interface of two cells, has been overhauled to make it
  more consistent with the rest of the library and use more intuitive names for functions. For example,
  \texttt{FEInterfaceValues::jump\_gradient()} is now \texttt{FEInterfaceValues::jump\_in\_shape\_gradients()}. Several new
  functions, such as \texttt{FEInterfaceValues::get\_jump\_in\_function\_values()}, have also been added.
  \item The \texttt{MeshWorker::ScratchData} and \texttt{MeshWorker::CopyData} have been made $hp$-compatible,
  and support face integration where the integration rule and mapping differs on either side of an interface.
  The \texttt{MeshWorker::CopyData} class has also been made compatible with complex numbers.
  \item Vectors attached to \texttt{DataOut} do not need to be in
    ghosted state anymore. Internally, we create a copy of the vector
    with appropriate ghost elements.
\end{itemize}
%
The changelog lists more than 100 other features and bugfixes.

@@

This release of \dealii contains a number of large and significant changes,
which will be discussed in this section.
It of course also includes a
vast number of smaller changes and added functionality; the details of these
can be found
\href{https://dealii.org/developer/doxygen/deal.II/changes_between_9_3_3_and_9_4_0.html}
{in the file that lists all changes for this release}; see \cite{changes94}.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Advances in simplex- and mixed-mesh support}\label{sec:simplex}

\dealii has supported simplex and mixed meshes since the previous
release, 9.3, on an experimental basis.
We have continued to work on this support; the current state is
sufficient to support several larger-scale programs, although not all
functionality in \dealii works with such meshes yet. Specifically,
we have fixed many bugs and generalized existing functions that
previously only worked for hypercube-shaped cells. The most notable
new functionalities are:

\begin{figure}

  \centering

  \phantom{.}
  \hfill
  \begin{tikzpicture}[scale=1.7]

    \coordinate (0) at (0,0);
    \coordinate (1) at (1,0);
    \coordinate (2) at (0,1);

    \coordinate (3) at (0,0.5);
    \coordinate (4) at (0.5,0.5);
    \coordinate (5) at (0.5,0);

    \foreach \i in {0,1, 2, 3, 4, 5}
    \draw[blue,fill=blue] (\i) circle (1.1pt) node [below] {};

    \draw (0) --(1) -- (2) -- (0);
    \draw (3) --(4) -- (5) -- (3);

  \end{tikzpicture}
  \hfill\hfill
  \begin{tikzpicture}[scale=1.3]

    \coordinate (0) at (0,0,0);
    \coordinate (1) at (0,0,1);
    \coordinate (2) at (1,0,0);
    \coordinate (3) at (0,1,0);
    \coordinate (4) at (0,0,0.5);
    \coordinate (6) at (0.5,0,0);
    \coordinate (7) at (0,0.5,0);
    \coordinate (5) at (0.5,0.0,0.5);
    \coordinate (8) at (0.0,0.5,0.5);
    \coordinate (9) at (0.5,0.5,0.0);

    \draw (0) -- (1) -- (2) -- (0);
    \draw (0) -- (1) -- (3) -- (0);
    \draw (1) -- (2) -- (3) -- (1);

    \draw[-, fill=red!30, opacity=.5] (4)--(6)--(7)--cycle;
    \draw[-, fill=red!30, opacity=.5] (0)--(6)--(7)--cycle;
    \draw[-, fill=red!30, opacity=.5] (0)--(4)--(7)--cycle;
    \draw[-, fill=red!30, opacity=.5] (0)--(4)--(6)--cycle;

    \draw[-, fill=blue!30, opacity=.5] (1)--(4)--(5)--cycle;
    \draw[-, fill=blue!30, opacity=.5] (1)--(4)--(8)--cycle;
    \draw[-, fill=blue!30, opacity=.5] (4)--(5)--(8)--cycle;
    \draw[-, fill=blue!30, opacity=.5] (1)--(5)--(8)--cycle;

    \draw[-, fill=green!30, opacity=.5] (2)--(5)--(6)--cycle;
    \draw[-, fill=green!30, opacity=.5] (2)--(6)--(9)--cycle;
    \draw[-, fill=green!30, opacity=.5] (2)--(5)--(9)--cycle;
    \draw[-, fill=green!30, opacity=.5] (5)--(6)--(9)--cycle;

    \draw[-, fill=orange!30, opacity=.5] (7)--(8)--(9)--cycle;
    \draw[-, fill=orange!30, opacity=.5] (3)--(7)--(9)--cycle;
    \draw[-, fill=orange!30, opacity=.5] (3)--(7)--(8)--cycle;
    \draw[-, fill=orange!30, opacity=.5] (3)--(8)--(9)--cycle;

    \foreach \i in {0,1,2,3,4,5,6,7,8,9}
    \draw[blue,fill=blue] (\i) circle (1.0pt) node [below] {};

  \end{tikzpicture}
  \begin{tikzpicture}[scale=1.3]

    \coordinate (0) at (0,0,0);
    \coordinate (1) at (0,0,1);
    \coordinate (2) at (1,0,0);
    \coordinate (3) at (0,1,0);
    \coordinate (4) at (0,0,0.5);
    \coordinate (6) at (0.5,0,0);
    \coordinate (7) at (0,0.5,0);
    \coordinate (5) at (0.5,0.0,0.5);
    \coordinate (8) at (0.0,0.5,0.5);
    \coordinate (9) at (0.5,0.5,0.0);

    \draw (0) -- (1) -- (2) -- (0);
    \draw (0) -- (1) -- (3) -- (0);
    \draw (1) -- (2) -- (3) -- (1);

    \draw[-, fill=red!30, opacity=.5] (4)--(5)--(6)--cycle;
    \draw[-, fill=red!30, opacity=.5] (4)--(5)--(8)--cycle;
    \draw[-, fill=red!30, opacity=.5] (4)--(6)--(8)--cycle;
    \draw[-, fill=red!30, opacity=.5] (5)--(6)--(8)--cycle;

%    \draw[-, fill=blue!30, opacity=.5] (4)--(6)--(7)--cycle;
%    \draw[-, fill=blue!30, opacity=.5] (4)--(6)--(8)--cycle;
%    \draw[-, fill=blue!30, opacity=.5] (4)--(7)--(8)--cycle;
%    \draw[-, fill=blue!30, opacity=.5] (6)--(7)--(8)--cycle;

%    \draw[-, fill=green!30, opacity=.5] (6)--(7)--(8)--cycle;
%    \draw[-, fill=green!30, opacity=.5] (6)--(7)--(9)--cycle;
%    \draw[-, fill=green!30, opacity=.5] (6)--(8)--(9)--cycle;
%    \draw[-, fill=green!30, opacity=.5] (7)--(8)--(9)--cycle;
%
    \draw[-, fill=blue!30, opacity=.5] (5)--(6)--(8)--cycle;
    \draw[-, fill=blue!30, opacity=.5] (5)--(6)--(9)--cycle;
    \draw[-, fill=blue!30, opacity=.5] (5)--(8)--(9)--cycle;
    \draw[-, fill=blue!30, opacity=.5] (6)--(8)--(9)--cycle;

    \foreach \i in {0,1,2,3,4,5,6,7,8,9}
    \draw[blue,fill=blue] (\i) circle (1.0pt) node [below] {};

  \end{tikzpicture}
  \begin{tikzpicture}[scale=1.3]

    \coordinate (0) at (0,0,0);
    \coordinate (1) at (0,0,1);
    \coordinate (2) at (1,0,0);
    \coordinate (3) at (0,1,0);
    \coordinate (4) at (0,0,0.5);
    \coordinate (6) at (0.5,0,0);
    \coordinate (7) at (0,0.5,0);
    \coordinate (5) at (0.5,0.0,0.5);
    \coordinate (8) at (0.0,0.5,0.5);
    \coordinate (9) at (0.5,0.5,0.0);

    \draw (0) -- (1) -- (2) -- (0);
    \draw (0) -- (1) -- (3) -- (0);
    \draw (1) -- (2) -- (3) -- (1);

%    \draw[-, fill=red!30, opacity=.5] (4)--(5)--(6)--cycle;
%    \draw[-, fill=red!30, opacity=.5] (4)--(5)--(8)--cycle;
%    \draw[-, fill=red!30, opacity=.5] (4)--(6)--(8)--cycle;
%    \draw[-, fill=red!30, opacity=.5] (5)--(6)--(8)--cycle;

    \draw[-, fill=red!30, opacity=.5] (4)--(6)--(7)--cycle;
    \draw[-, fill=red!30, opacity=.5] (4)--(6)--(8)--cycle;
    \draw[-, fill=red!30, opacity=.5] (4)--(7)--(8)--cycle;
    \draw[-, fill=red!30, opacity=.5] (6)--(7)--(8)--cycle;

    \draw[-, fill=blue!30, opacity=.5] (6)--(7)--(8)--cycle;
    \draw[-, fill=blue!30, opacity=.5] (6)--(7)--(9)--cycle;
    \draw[-, fill=blue!30, opacity=.5] (6)--(8)--(9)--cycle;
    \draw[-, fill=blue!30, opacity=.5] (7)--(8)--(9)--cycle;

%    \draw[-, fill=orange!30, opacity=.5] (5)--(6)--(8)--cycle;
%    \draw[-, fill=orange!30, opacity=.5] (5)--(6)--(9)--cycle;
%    \draw[-, fill=orange!30, opacity=.5] (5)--(8)--(9)--cycle;
%    \draw[-, fill=orange!30, opacity=.5] (6)--(8)--(9)--cycle;

    \foreach \i in {0,1,2,3,4,5,6,7,8,9}
    \draw[blue,fill=blue] (\i) circle (1.0pt) node [below] {};

  \end{tikzpicture}
  \hfill
  \phantom{.}

  % We definitely supported tri refinement in 9.3 - see 6ceb8b70559dc2274761a28dc7836f1e9602da8c (Dec 4, 2020)
  % and also 74b8171dd030a4b2de41955b8d913cde06974b91 (Feb 4, 2021)
  \caption{\it The previous release, 9.3 \cite{dealII93},
    added support for adaptive mesh 
  refinement with triangles. This release adds support for global
  tetrahedral refinement by subdividing each tetrahedron into eight
  children, as shown on the right.}
  \label{fig:refinement}
\end{figure}

\begin{itemize}
\item Experimental support for locally refined meshes: For finite
  element computations on locally refined
meshes, one needs (i) the possibility to locally refine the mesh (see
Figure~\ref{fig:refinement}), and (ii) appropriate hanging-node
constraints. Both are now available in 2D; for 3D, the implementation
of the constraint definitions is still in progress.
\item \texttt{QIteratedSimplex} allows to build composite simplex quadrature rules.
\item The new wrappers to the \texttt{CGAL} library allow for the easy
  creation of simplex
meshes. For more details, see Section~\ref{sec:cgalwrappers}.
\end{itemize}

Furthermore, we have continued to remove uses of the
\texttt{GeometryInfo} class (which is specific to hypercube cells) from
the library, and to replace them with more general equivalent
functionality based on the
\texttt{ReferenceCell} class. Once all instances of
\texttt{GeometryInfo} are removed, we will deprecate the class.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Repartitioning of distributed meshes}\label{sec:repartitioning}

\dealii has two classes that support distributed storage of meshes:
\texttt{parallel::\allowbreak distributed::\allowbreak Triangulation}
(which in the following we will abbreviate as \texttt{p::d::T}) and
\texttt{parallel::\allowbreak fully\allowbreak distributed::\allowbreak
  Triangulation} (or, in short, \texttt{p::f::T}). The former is
partitioned based on the space-filling Morton (or Z-order) curve
implemented by the \pfrst backend. The latter, more recent
class is currently statically partitioned at creation time.

There are now new utility functions in the
\texttt{RepartitioningPolicyTools} namespace that can be
used to create a new \texttt{p::f::T} instance,
given a distributed triangulation (\texttt{p::f::T} or
\texttt{p::d::T}) and a vector that indicates the designated owner processes of locally
owned cells. This approach allows for partitioning \texttt{p::f::T}
based on arbitrary criteria. The workflow is shown in the following listing:
\begin{c++}
// Original triangulation:
parallel::distributed::Triangulation<dim> tria(communicator);

// Select a partitioning policy and use it:
const RepartitioningPolicyTools::CellWeightPolicy<dim> policy(tria, fu);
const auto construction_data = TriangulationDescription::Utilities::
  create_description_from_triangulation(tria, policy.partition(tria));

// Create the new triangulation:
parallel::fullydistributed::Triangulation<dim> tria_pft(comm);
tria_pft.create_triangulation(construction_data);
\end{c++}

\begin{figure}
    \centering
    \def\svgwidth{0.8\columnwidth}
    \input{svg/repartitioning.pdf_tex}
    \caption{\it Visualization of the repartitioning process and the
      construction of the new mesh of process 0. Top
      left: Existing ownership of the cells of the mesh, distributed
      on four processes. Bottom left:
      Requested new ownership. Second column: Processes 0 and 1
      collect information about the cells they own or that are ghost
      cells. Processes 2 and 3 do not contribute to process 0. Third column: What
      processes 0 and 1 would send to process 0. Fourth column: The
      combined knowledge on process 0.}\label{fig:repartitioning}
\end{figure}

The setup process (of \texttt{construction\_data}) is visualized in Figure~\ref{fig:repartitioning}. At
first, locally owned cells and their
surrounding (ghost) cells are collected on each process and
sent to the new owner. On the
receiving side, the sets of all cells are combined and possible duplicates
are removed. This information is enough to construct a new triangulation.
For sending/receiving, we apply consensus-based algorithms~\cite{hoefler2010scalable}, which
we introduced into the library in release 9.2~\cite{dealII92} -- see
also Section~\ref{sec:CA}. This kind of consensus-based
algorithm is used also in \cite{ibanez2016pumi} for repartitioning.

In addition to the predefined partitioning policies, users
can write their own by implementing the
\texttt{RepartitioningPolicyTools::Base} interface. Like the
``active'' level of finest cells, the scheme also allows for
arbitrarily repartitioned multigrid levels.

In future releases, we plan to add support for repartitioning based on distributed graph
partitioning libraries, e.g., \texttt{ParMETIS} or \texttt{Zoltan}. Furthermore, we intend to extend
\texttt{p::f::T} to support adaptive mesh refinement, completing \texttt{p::f::T}
as a complement to \texttt{p::d::T}.

@@

The matrix-free infrastructure of \dealii is used to solve problems
without assembling matrices, providing only the \textit{action} of the matrix
instead. It has seen a number of new features in this release, the most
notable ones being:
\begin{itemize}
\item Improved support for computations with Hessians of shape functions: Just like values and gradients, Hessians can be
now evaluated and integrated during matrix-free loops
both for cells and faces. This could enable, for example, writing
a matrix-free version of step-47, which solves the biharmonic equation with the discontinuous Galerkin method.
\item Cell-centric loops now also allow access to gradients and Hessians
of neighboring cells on faces. The major difficulty here was the relative orientation
of cells on unstructured meshes.
\item Users can now create their own cell batches, by providing \texttt{FEEvaluation::reinit()} a list of cell IDs. \texttt{FEEvaluation}
accesses the appropriate data and reshuffles mapping data accordingly on
the fly in order to enable vectorization over cells. The new feature is useful in several
contexts, for example for sharp interfaces (e.g., two-phase flow
or shock capturing), where one needs to treat cells that are ``cut'' by
the interface in a special way. A challenge is that cell batches
might contain cut or non-cut cells, effectively counteracting the cross-cell vectorization.
Previous functionality has provided the option of masking certain cells in cell batches, which works well if
the code paths do not diverge too much, or to
categorize cells during \texttt{MatrixFree::reinit()} in such a way that mixed cell batches do not occur. However, \texttt{MatrixFree::reinit()}
might be too expensive if recategorization needs to happen very frequently to follow the dynamics of a system, e.g., in each time step. Despite some overhead compared to static matrix-free loops, the new feature can be the best option in such dynamic scenarios.
\item Initial support for H(div)-conforming elements with Piola transform, based on Raviart--Thomas finite element spaces with the updated class \texttt{FE\_RaviartThomasNodal}, has been added. This feature is currently limited to meshes in standard orientation and affine geometries. Full support and performance optimizations will be provided in a future release.
\item Selected matrix-free algorithms can now exploit additional data locality
  between the matrix-vector product and vector operations happening nearby in
  an algorithm. Interfaces have been added to both the
  \texttt{PreconditionChebyshev} class (a frequently used smoother in multigrid
  methods) and the conjugate gradient implementation in \texttt{SolverCG}.
  From the user's perspective, an operator needs to define a \texttt{vmult} operation,
  taking two additional \texttt{std::function} arguments. The first function
  defines the operation to be scheduled on the vector entries before the
  matrix-vector product touches them, and the second what happens
  afterward. The new features also include a renumbering to maximize data
  locality. The theory is described in the
  contribution~\cite{kronbichler2022cg}.
\end{itemize}
Besides these new features, we improved the performance of
hanging-node-constraint evaluation on the CPU. Instead of performing
quasi-dense matrix-vector multiplications~\cite{KronbichlerKormann2012}, we now use an
approach based on in-place interpolation and sum factorization, similar
to what was already done in the GPU code~\cite{ljungkvist2017matrix}. In \cite{munch2022hn}, the algorithm is described
and performance numbers are shown, indicating a reduction
of overhead of cells with hanging nodes by a factor of ten.

Finally, we have performed a major restructuring of the internals
of the \texttt{FEEvaluation} classes. This reduces some overhead for low polynomial degrees and will enable us to add support for new element types in the future.

We would like to remind users that we transitioned from the use of Booleans
to flags to configure the evaluation and integration process of \texttt{FEEvaluation}
and \texttt{FEFaceEvaluation}:

\begin{c++}
fe_eval.evaluate(false, true, false)         // old (deprecated)
fe_eval.evaluate(EvaluationFlags::gradients) // new
\end{c++}

%\begin{c++}
%additional_data.mapping_update_flags = ... | update_hessians;
%\end{c++}
%
%\begin{c++}
%FEEvaluation<dim> phi(matrix_free);
%
%phi.reinit(cell);
%phi.gather_evaluate(src, ... | EvaluationFlags::hessians);
%
%for(const auto q : phi.quadrature_point_indices())
%  phi.submit_hessian(phi.get_hessian(q), q);
%
%phi.integrate_scatter(... | EvaluationFlags::hessians, dst);
%\end{c++}
%
%... similar for faces; can be used to implement biharmonic equation (like in
%step-X)

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Advances in multigrid infrastructure}\label{sec:multigrid}

In release 9.3~\cite{dealII93}, we added support for global-coarsening multigrid in
addition to the established local-smoothing infrastructure. Global
coarsening algorithms smoothen over the whole computational domain on each
multigrid level, which is obtained by coarsening the finest cells of
the next finer multigrid level.
For this purpose, we use a sequence of triangulations, and we perform
the smoothing only on their active levels. To create the sequence of
triangulations, one can use the functions \texttt{MGTransferGlobalCoarseningTools::create\_geometric\_coarsening\_sequence()}. A new version takes an
instance of \texttt{RepartitioningPolicyTools::Base} (see Subsection~\ref{sec:repartitioning}) as argument, which allows specifying the parallel
distribution of each multigrid level (in contrast to the fixed first-child policy in the case of local smoothing). These features have been developed and tuned for running on a supercomputer scale
with complicated coarse meshes as presented in \cite{kronbichler2021next}.
Furthermore, we added support for block vectors,
fixed a number of limitations, and performed performance optimizations of
the transfer operator; particularly, the redundant copy from/to temporary vectors
has been eliminated. Furthermore, hanging-node constraints are applied
efficiently in the same way as in the matrix-free loops (see Subsection~\ref{sec:mf}).

In \cite{munch2022gc}, the performance of the local-smoothing and global-coarsening
infrastructure of \dealii was compared for locally refined meshes. The results indicate that
the local definition of multigrid levels might introduce load imbalances
in the case of local smoothing so that global coarsening is favorable despite
potentially more expensive intergrid transfers. In order to assess the benefits
of one approach against the other, \dealii provides new functions
\texttt{workload\_imbalance()} and \texttt{vertical\_communication\_efficiency()}
in the \texttt{MGTools} namespace for the  estimation of the imbalance during, e.g.,
smoothing or the
communication efficiency during intergrid transfer, purely based on the given mesh.
% MK: I would not add this part, it does not really fit into this paper as
% there is nothing to report at this point (and we should then add references
% to the actual literature).
%The publication \cite{munch2022gc} also points out that not
%all types of smoothers are applicable for global coarsening due to the
%presence of hanging nodes, which is a motivation to add new smoother types
%to \dealii in the future.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{CutFEM support}\label{sec:cut}

Several classes have been added to the \texttt{NonMatching} namespace to enable the use of cut finite element methods~\cite{burman_cutfem_2015}.
In the literature, these types of methods are also referred to as immersed, extended, or fictitious finite element methods.
Here, the domain, $\Omega$, is immersed in the background mesh, as illustrated in Fig.~\ref{fig:immersed-domain}.
Often, one solves for the degrees of freedom of the smallest submesh
which completely covers the domain, i.e., the blue and yellow cells of
Fig.~\ref{fig:location-to-level-set}.
The bilinear form in the weak form would then, for example, look like
\begin{equation}
  a(u,v) = (\nabla u, \nabla v)_\Omega - (\partial_n u, v)_\Gamma + \ldots
\end{equation}
Thus, when assembling on a cut cell $K$, we are
required to integrate over the part of the domain and the part of the boundary, $\Gamma = \partial \Omega$, that falls inside the cell:
$K\cap \Omega$ and $K \cap \Gamma$.
Many of the new classes that support these operations assume that the domain is described by a level set function,
$\psi : \mathbb{R}^d \to \mathbb{R}$, such that
\begin{align}\label{eq:levelset}
  \Omega = \{x \in \mathbb{R}^d : \psi(x)<0\},
  \qquad
  \Gamma = \{x \in \mathbb{R}^d : \psi(x) = 0\}.
\end{align}

\begin{figure}
  \centering
  \begin{subfigure}[b]{0.3\textwidth}
    \centering
    \includegraphics[height=.12\paperheight]{svg/immersed-domain.pdf}
    \caption{\label{fig:immersed-domain}}
  \end{subfigure}
  \qquad
  \begin{subfigure}[b]{0.3\textwidth}
    \centering
    \includegraphics[height=.12\paperheight]{svg/location-to-level-set.pdf}
    \caption{ \label{fig:location-to-level-set}}
  \end{subfigure}
  \caption{\it (a) Domain immersed in a background mesh. (b) Value of \texttt{LocationToLevelSet} for each cell.}
%\end{figure}
%
%\begin{figure}
  \centering
  \includegraphics[width=.4\paperwidth]{svg/immersed_quadratures.pdf}
  \caption{\it Quadrature points for integrating over the three different regions of a cell cut by the zero contour of the level set function, $\psi$. \label{fig:immersed_quadratures}}
\end{figure}

Specifically, the following are the key new classes and functions:
\begin{itemize}
  \item The \texttt{MeshClassifier} class identifies how the active cells and faces are located relative to the zero contour of the level set function, as illustrated in Figure~\ref{fig:location-to-level-set}. Its member function \texttt{location\_to\_level\_set()} takes a cell or face and
        returns an enum, \texttt{LocationToLevelSet}, with values \{\texttt{inside}, \texttt{outside}, \texttt{intersected}\}.
        This information is typically needed when choosing what element (e.g., \texttt{FE\_Q} or \texttt{FE\_Nothing}) and/or what quadrature (defined over the 
        complete cell or over a part---see below) a cell
        should use.

>>>>

\hypersetup{
  pdfauthor={
    Daniel Arndt,
    Wolfgang Bangerth,
    Marc Fehling,
    Timo Heister,
    Luca Heltai,
    Martin Kronbichler,
    Matthias Maier,
    Peter Munch,
    Jean-Paul Pelteret,
    Bruno Turcksin,
    David Wells
  },
  pdftitle={The deal.II Library, Version 9.5, 2023},
}

\title{The \dealii Library, Version 9.5}

@@

@@

\author[4]{Timo~Heister}
 \affil[4]{School of Mathematical and Statistical Sciences,
   Clemson University,
   Clemson, SC, 29634, USA
   {\texttt{heister@clemson.edu}}}

\author[5]{Luca~Heltai}
\affil[5]{SISSA,
   International School for Advanced Studies,
   Via Bonomea 265,
   34136, Trieste, Italy.
   {\texttt{marco.feder/luca.heltai@sissa.it}}}

 \author[6]{Martin~Kronbichler}
 \affil[6]{Institute of Mathematics,
   University of Augsburg,
   Universit\"atsstr.~12a, 86159 Augsburg, Germany.
   {\texttt{martin.kronbichler@uni-a.de}}}

\author[7]{Matthias~Maier}
\affil[7]{Department of Mathematics,
  Texas A\&M University,
  3368 TAMU,
  College Station, TX 77845, USA.
  {\texttt{maier@math.tamu.edu}}}

\author[6,8]{Peter Munch}
 \affil[8]{Institute of Material Systems Modeling,
 Helmholtz-Zentrum Hereon,
 Max-Planck-Str. 1, 21502 Geesthacht, Germany.
   {\texttt{peter.muench@hereon.de}}}

\author[9]{Jean-Paul~Pelteret}
\affil[9]{Independent researcher.
{\texttt{jppelteret@gmail.com}}}

@@

\author[10]{David Wells}
\affil[10]{Department of Mathematics, University of North Carolina,
  Chapel Hill, NC 27516, USA.
  {\texttt{drwells@email.unc.edu}}}

@@

\begin{abstract}
  This paper provides an overview of the new features of the finite element
  library \dealii, version 9.5.
\end{abstract}

@@

\dealii version 9.5.0 was released XX XX, 2023.
This paper provides an
overview of the new features of this release and serves as a citable
reference for the \dealii software library version 9.5. \dealii is an
object-oriented finite element library used around the world in the
development of finite element solvers. It is available for free under the
GNU Lesser General Public License (LGPL). Downloads are available at
\url{https://www.dealii.org/} and \url{https://github.com/dealii/dealii}.

The major changes of this release are:
%
\begin{itemize}
  \item Integration of \texttt{Kokkos} (see Section~\ref{sec:kokkos});
  \item Update of the \texttt{PETSc} infrastructure (see Section~\ref{sec:petsc});
  \item Addition of new \texttt{Trilinos} wrappers (see Section~\ref{sec:trilinos});
  \item Advances in matrix-free infrastructure (see Section~\ref{sec:mf});
  \item Advance in non-matching support (see Section~\ref{sec:nonmatching});
  \item New liner algebra features (see Section~\ref{sec:lac}).
\end{itemize}
%

While all of these major changes are discussed in detail in
Section~\ref{sec:major}, there
are a number of other noteworthy changes in the current \dealii release,
which we briefly outline in the remainder of this section:
%
\begin{itemize}
  \item The \texttt{CellAccessor::as\_dof\_handler\_iterator()}
\end{itemize}
%
The changelog lists more than X other features and bugfixes.

@@

This release of \dealii contains a number of large and significant changes,
which will be discussed in this section.
It of course also includes a
vast number of smaller changes and added functionality; the details of these
can be found
\href{https://dealii.org/developer/doxygen/deal.II/changes_between_9_4_2_and_9_5_0.html}
{in the file that lists all changes for this release}; see \cite{changes95}.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Integration of Kokkos}\label{sec:kokkos}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Update of the PETSc infrastructure}\label{sec:petsc}

\begin{itemize}
\item PreconditionBDDC
\end{itemize}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{New Trilinos wrappers}\label{sec:trilinos}

\begin{itemize}
\item Belos
\item NOX
\end{itemize}

@@

\begin{itemize}
\item element activation and deactivation, \texttt{FE\_Nothing}, categorization, DG, face integrals 
\item pre/post for DG
\item \texttt{CellwiseInverseMassMatrix} operator for coupling (dyadic) coefficients
\end{itemize}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Advance in non-matching support}\label{sec:nonmatching}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{New linear algebra features}\label{sec:lac}

\begin{itemize}
  \item \texttt{SolverGMRES}: classical/modified Gram Schmidt process
  \item James Lottes’s fourth-kind Chebyshev
  \item \texttt{PreconditionChebyshev}, \texttt{PreconditionRelaxation}
  \item restrict matrices
  \item fast diagonalization method
\end{itemize}
